---
title: "EDA Project"
author: "Alexander Rodriguez"
date: "1/19/2018"
output: html_document
---

Prosper Loan Data by Alexander Rodriguez
========================================================

> **Introduction**: Credit is an ubiquitous aspect of our modern day lives.
Credit allows people to borrow money for life expenses that one cannot afford
without a loan.  It is important to maintain a good credit rating 
to ensure that loan terms are favorable to the borrower.  This project 
will provide visualizations about the Prosper Loans Dataset and the credit 
terms that borrowers receive based on their credit history.
The Prosper Loans dataset contains information on custumer credit rating, 
loan status, loan amount and many other variables that would help to 
understand what factors are considered when credit is approved. Below are 
the two over-arching questions which the EDA process should reveal answers to.

> **Questions**: What kinds of loan terms do people with excellent, good,
fair, poor, and bad credit receive? What are the advantages to having a good
credit score?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
loan <- read.csv('prosperLoanData.csv', header=T, na.strings=c(""," ","NA"))
```

# Univariate Plots Section

# 1. Check the credit rating data
```{r echo=FALSE,  message=FALSE, warning=FALSE, Univariate_Plots}
# check the credit rating data 
summary(loan$prosper_rating_a)

#convert the credit_grade column to characters
loan$credit_grade <- as.character(loan$credit_grade)

#convert the credit_grade column to characters
loan$prosper_rating_a <- as.character(loan$prosper_rating_a)

#fill in the prosper_rating missing data with the values in credit_grade to 
#make a complete prosper_rating column
loan$prosper_rating_a[is.na(loan$prosper_rating_a)]<- 
  loan$credit_grade[is.na(loan$prosper_rating_a)]

#convert prosper_rating_a data to ordered factors
loan$prosper_rating_a <-factor(loan$prosper_rating_a, 
                               levels = c('AA', 'A', 'B', 'C', 
                                          'D', 'E', 'HR', 'NC' ), 
                               ordered=TRUE)

#create a bar graph with proper_rating(credit grade)
ggplot(aes(prosper_rating_a), data=loan) +
         geom_bar()

```

***
> **Comment**: The prosper_rating_a column missing values were
filled in with a corresponding credit score which will greatly 
help the analysis.

> **Observation**: The mode for prosper_rating_a is a rating of 'C'.  
Also,  many people with bad and poor credit recieve loans.
***

# 2. Investigate borrower APR
```{r echo=FALSE, message=FALSE, warning=FALSE}

# summary of borrower APR by credit rating
by(loan$borrower_apr, loan$prosper_rating_a, summary)

#display a histogram of the borrower apr 
ggplot(aes(borrower_apr), data=loan) + 
  geom_histogram(color='black', fill ='#099DD9') 
 

# histogram of the borrower_apr with a facet_wrap of prosper_rating_a
ggplot(aes(borrower_apr), data=loan) + 
  geom_histogram(color='black', fill ='#099DD9') +
  facet_wrap(~prosper_rating_a)
```

***
> **Observation**: The distribution of borrower apr appears to 
be normally distributed.

> **Observation**: To provide some insight on the first question of 
"What are the advantages of having good credit?", from the summary 
the Max APR is the lowest for the highest credit rating of 'AA'. 
It follows that as the credit rating decreases the Max apr increases.  
Furthermore,  the mean apr follows this trend as well.  
The faceted histograms are also included as visualizations.

***

# 3. Investigate loan status
```{r echo=FALSE,  message=FALSE, warning=FALSE}
# check the loan status
ggplot(aes(loan_status), data=loan) +
         geom_bar() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Proportion of borrowers who defaulted
1 - length(which(loan$loan_status == 'Completed' | 
                   loan$loan_status == 'Current' )) / length(loan$loan_status)

length(which(loan$loan_status == 'Chargedoff' )) / length(loan$loan_status)
```

***
> **Observation**: The majority of the loans are either current or have been
completed, 16.9% defaulted on their loan(This includes Charge-offs which
is equal to 10.5%). It would be interesting to determine what credit rating
has the highest proportion in the chargedoff and defaulted categories.

***

#4. Investigate loan term
```{r echo=FALSE,  message=FALSE, warning=FALSE}

# boxplot of loan term
ggplot(aes(x= term), data = loan) +
  geom_bar() +
  scale_x_continuous(breaks = c(12, 36, 60))

#boxplot of loan term
ggplot(aes(x= term), data = loan) +
  geom_bar() +
  facet_wrap(~prosper_rating_a) +
  scale_x_continuous(breaks = c(12, 36, 60))
```

***
> *Observation*: Most people get a loan term of 36 months.  
Are the longer term loans for larger amount of money borrowed?

***

# 5. Investigate debt to income ratio. 
```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(DebtToIncomeRatio), data = loan) +
  geom_histogram()

# histogram of debt to income ratio
ggplot(aes(DebtToIncomeRatio), data = loan) +
  geom_histogram() + 
  scale_x_continuous(name="Debt to Income Ratio", 
                     breaks = seq(0, 1, .1),
                     limits=c(0,1) ) +
  facet_wrap(~prosper_rating_a, scales = 'free')

# Determine the percentage of borrowers who have a debt to income ratio
# less than 30%.
length(which(loan$DebtToIncomeRatio <= 0.30)) / length(loan$DebtToIncomeRatio)

summary(loan$DebtToIncomeRatio)
```

***
> *Observation*: 67.4% of borrowers have debt to income ratio that is 30% or
less. What factors are the most significant to be approved for a loan when
a person has more than 30% debt to income ratio? Why are people being
approved for loans with debt to income ratios greater than 100%?

***

# 6. Investigate Monthly Loan Payment 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
by(loan$MonthlyLoanPayment, loan$prosper_rating_a, summary)

ggplot(aes(MonthlyLoanPayment), data = loan) +
  geom_histogram() 

ggplot(aes(x=prosper_rating_a, y= MonthlyLoanPayment), data = loan) +
  geom_boxplot()
```

*** 
> *Observation*: Borrowers with AA, A, and B have very similar distributions
and the data shows that those borrowers are trusted with higher loan amounts
and thus higher montly payments.

***

# 7. Investigate employment status and employment duration.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(EmploymentStatus), data =loan) +
  geom_bar() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))

summary(loan$employment_status_duration)

ggplot(aes(employment_status_duration), data = loan) +
  geom_histogram() +
  scale_x_sqrt()
```

***
> *Observation*: As expected most people are employed, however, some people are
not employed.  I wonder how they were able to get a loan.  The distribution
of employment_status_duaration is positively skewed and appears to be
normally distributed with the square root transformation.

***

#8. Univariate Analysis

> **Summary**: This data set is replete with 81 features. 
One of the main features that I chose to investigate was credit rating which
was titled'prosper_rating_a'.  Credit rating is an interesting feature to study 
since it is a summary of a borrower's credit worthiness.  Through out the 
various plots that I created, there is an obvious advantage to having a credit
rating that is 'AA', 'A', and 'B', since these credit ratings are connected to
lower aprand higher loan amounts.  
Other features of interest are debt-to-income ratio which ideally should be less
than 30%, however, for some borrowers the ratio was above 100% and some as much 
as 1000%.  I would like to understand why a person would get granted a loan 
with such a high debt-to-income ratio.  
One feature in the dataset that will help support the investigation would be 
the actual credit score.  Especially how the specific value of a credit score
can affect credit terms.


# Bivariate Plots Section

# 9. Investigate the relationship between Prosper Score and borrower apr
```{r echo=FALSE,  message=FALSE, warning=FALSE, Bivariate_Plots}

#A custom risk score built using historical Prosper data.
#The score ranges from 1-10, with 10 being the best, or lowest risk score.  
#Applicable for loans originated after July 2009.
summary(loan$ProsperScore)

# the description of the ProsperScore indicates that a score of 10
# is the highest.  Thus, any score of 11 will be treated as a mistake
# and changed to a 10.

# Find the rows with ProsperScore of 11.
ind <- which(loan$ProsperScore == 11)

#change 11 to 10.
loan$ProsperScore[ind] <- 10

ggplot(aes(x= ProsperScore, y = borrower_apr ), data=loan) +
  
  geom_jitter(alpha = 1/20)
  scale_x_continuous(breaks = seq(0,10, 1),
                     limits = c(1, 10))
```

***
> *Observation*: There is big advantage to having a prosper score of 10 vs. 1.
However, there seems to be very little advantage of having a 7,6 or a 5.

***
  
# 10A. Investigate borrower_apr vs. prosper score.
```{r echo=FALSE,  message=FALSE, warning=FALSE}
summary(loan$CreditScoreRangeLower)
summary(loan$borrower_apr)
#display how the relationship between the credit score and the borrower_apr
p1 <- ggplot(aes(x=CreditScoreRangeLower, y=borrower_apr), data=loan) +
      geom_point(alpha = 1/10, position=position_jitter(h=0)) +
      scale_x_continuous(limits = c(400, 880))

p2 <- ggplot(aes(x= ProsperScore, y = borrower_apr ), data=loan) +
     
      geom_jitter(alpha=1/20)
     scale_x_discrete(breaks = seq(0,10, 1),
                      limits = c(1, 10))

grid.arrange(p1, p2, ncol =1)
```

***
> *Observation*: There is a negative relationship between borrower_apr
and CreditScoreRange. However, there is a lot of variability with
people with high credit scores (above 700) receiving a loan with
an apr that is above the average APR ofabout 20%.

***

#10B. Look at the average borrower apr vs credit score\
to look at the trend.
```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x=CreditScoreRangeLower, y= borrower_apr), data=loan) +
 
   scale_x_continuous(limits = c(400, 880)) +
   scale_y_continuous(name="Borrower APR",
                     breaks = seq(0, .3, .1), limit = c(0,.3)) +
   geom_line(stat='summary', fun.y=mean)  
   

ggplot(aes(x=ProsperScore, y= borrower_apr), data=loan) +
  scale_y_continuous(name="Borrower APR",
                     breaks = seq(.1, .3, .1), limit = c(.1,.3)) +
  geom_line(stat='summary', fun.y=mean)
```

***
> *Observation*: The average borrower apr is clearly has a negative relationship
with prosper and credit score.  The higher the prosper score the lower the 
average apr.

***

#11. Does DebtToIncomeRatio help to predict ProsperScore?
```{r echo=FALSE,  message=FALSE, warning=FALSE}
loan$DebtToIncome.bucket <- cut(loan$DebtToIncomeRatio, 
                                breaks=c(0,0.14,0.22,0.32,0.60,10.01))
ggplot(aes(x = DebtToIncome.bucket, y = ProsperScore), 
       data= subset(loan, !is.na(ProsperScore))) +
  geom_boxplot(color = 'black', fill = '#099DD9') +
  ggtitle('Prosper Score in groups of Debt to Income') 
```

***
> *Observation*: The lowest debt-to-income buckets have the highest median
Prosper Score.  Meaning that as the debt-to-income ratio increases a 
borrowers scores is expected to decrease.

***

# 12. Does Credit Score help predict ProsperScore?
```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x = CreditScoreRangeLower, y= ProsperScore), data=loan) +
  
      geom_bin2d() +
      scale_x_continuous(limits = c(550, 880))+
      scale_y_continuous(name="Prosper Score",
                     breaks = seq(0, 10, 1)) +
      geom_smooth(method='lm', color='red')
```

***
> *Observation*: Having a good credit(above 750) score does not guarantee a 
good prosper score (8,9 or 10).  In fact, there are some borrowers with
a Credit Score above 800 that have a Prosper Score below 4.  

***

#13. Determine the outliers in the Debt to income ratio category
```{r echo=FALSE,message=FALSE, warning=FALSE}
OutVals = boxplot(loan$DebtToIncomeRatio, plot=FALSE)$out
#which(loan$DebtToIncomeRatio %in% OutVals)
summary(OutVals)
```

***
> *Observation*: The minimum outlier is a Debt to income ratio of 60%.

***

# 14. Investigate credit score and debt to income ratio
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x=CreditScoreRangeLower, y=DebtToIncomeRatio), data=loan) +
   geom_point(alpha = 1/4, position=position_jitter(h=0), 
              color = 'orange', fill = 'black') +
   scale_x_continuous(limits = c(400, 880)) +
   scale_y_continuous(name="Debt to Income Ratio",
                     breaks = seq(0, 1, .1), limit = c(0,1))
```

***
> *Observation*: Borrowers with Credit scores that are lower than
600 can receive a loan but the data shows that their debt to income
ratios must be lower than borrowers with higher scores.

***

# 15. Investigate Debt to income ratio and loan status.
```{r echo=FALSE,  message=FALSE, warning=FALSE}

ggplot(aes(x=DebtToIncome.bucket), data=loan) +
  geom_bar(aes(fill=loan_status)) 
```

***
> *Comment*: A DebtToIncome.bucket variable was created by using the 
five number summary. The minimum outlier is 0.06 and the maximum value
in the data is 10.01.

> *Observation*: Most people either pay off their loan or are in good
standing.

# 16. Investigate employment_status_duration vs. borrower apr
```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x= employment_status_duration, y= borrower_apr),
       data = loan) +
  geom_point(alpha = 1/2, position=position_jitter(h=0)) 

```

***
> *Observation*: Having a long emplyment status duration does not
guarantee a low borrower apr.

***

#17. Investigate listing category and Homeownership
```{r echo=FALSE,  message=FALSE, warning=FALSE}
#The category of the listing that the borrower selected when posting their
#listing: 0 - Not Available, 1 - Debt Consolidation, 2 - Home Improvement,
#3 - Business, 4 - Personal Loan, 5 - Student Use, 6 - Auto, 7- Other, 
#8 - Baby&Adoption, 9 - Boat, 10 - Cosmetic Procedure, 11 - Engagement Ring,
#12 - Green Loans, 13 - Household Expenses, 14 - Large Purchases,
#15 - Medical/Dental, 16 - Motorcycle, 17 - RV, 18 - Taxes,
#19 - Vacation, 20 - Wedding Loans

table(loan$IsBorrowerHomeowner )
ggplot(aes(x=listing_category_n), data=loan) +
  geom_bar(aes(fill=IsBorrowerHomeowner)) +
  coord_cartesian(xlim = c(0,20))                     
```

***
> *Observation*: Roughly 50% of borrowers are homeowners.  The majority of 
loans are for debt consolidation.

***

# 18. Bivariate Analysis

> **Summary**: In this section I investigated the relationship between both
prosper score and credit score with borrower apr.  There is a clear advantage
to having a high credit score and a high prosper score to getting an borrower 
apr that is below the 20% average.  However, there are borrowers who received 
an apr that is above the average and have a high prosper score(8,9, 10).  
Perhaps, there are other factors that have an effect on borrower apr such as 
occupation and employment status duration.
Something of interest arose when I looked at employment status duration vs 
borrower apr.  There is no observable relationship between having a long
employment status duration and receiving a favorable borrower apr.
The strongest relationship that I found in this section was the borrower
debt-to-income-ratio vs. credit score range.  The data indicates that 
borrowers with credit scores above 800 have lower debt-to-income ratios 
and borrowers who have scores less that 500 
must have a debt-to-income-ratio that is less than the average.

# Multivariate Plots Section

# 19. Investigate Monthly Loan Payment and Customer Principal Payments

```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x=MonthlyLoanPayment, y= LP_CustomerPrincipalPayments, 
       color = DebtToIncome.bucket), 
       data=loan) +
   geom_point() + 
  scale_y_sqrt() +
  scale_color_brewer(type='div')
```

***
> *Observation*: I notice that borrowers with the lowest deb to income
ratio have montly payment compared to their loan amount.

***

# 20. Investigate prosper score, borrower apr and debt-to-income-ratio
```{r echo=FALSE,  message=FALSE, warning=FALSE, Multivariate_Plots}
loan$DebtToIncome.bucket <- cut(loan$DebtToIncomeRatio, 
                                breaks=c(0,0.14,0.22,0.32,0.60,10.01))

ggplot(aes(x= ProsperScore, y = borrower_apr, color = DebtToIncome.bucket ), data=loan) +

  geom_jitter() +
  scale_x_continuous(breaks = seq(0,11, 1),
                     limits = c(1, 10)) +
  scale_color_brewer(type='div')  

# Count how many borrowers receive 10% or lower apr and have a high 
# debt to income
low_apr <- which(loan$borrower_apr <= 0.10 & loan$DebtToIncome.bucket == '(0.32,0.6]')
length(low_apr)

```

***
> *Observation*: This plot shows that to receive an borrower apr below 10% a 
borrower must be in the lowest DebtToIncome bucket(0,0.14].  However, 
some borrowers(total of 303) in the (0.32, 0.60] bucket also received
an borrower apr of 10%.

***

# 21. Loan status and Monthly payment colored by prosper rating
```{r echo=FALSE,  message=FALSE, warning=FALSE}
p1 <- ggplot(aes(x= loan_status, y = MonthlyLoanPayment, 
                 color=prosper_rating_a ), data=loan) +
   
      geom_jitter() +
      scale_color_brewer(type='qual') +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(aes(x= loan_status, y = MonthlyLoanPayment ), data=loan) +
      geom_boxplot(color='black', fill ='#099DD9') +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
      
grid.arrange(p1, p2, ncol =1)

by(loan$prosper_rating_a, loan$loan_status, summary)
```

***
> *Observation*: There is alot of blue and red colored dots in the defaulted
categories.  Conversely, there is a lot of green and violet colored dots 
in the completed and current categories.  Lastly, there are cases when
borrowers with good credit default on their loans and even get a chargeoff.

*** 

#22A.How does credit score and debt to income ratio affect borrower apr
```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x=CreditScoreRangeLower, y= borrower_apr, 
           color=DebtToIncome.bucket),
          data=loan) +
  geom_point(position=position_jitter(h=0)) +
   scale_x_continuous(limits = c(400, 880)) +
   scale_y_continuous(name="Borrower APR",
                     breaks = seq(0, .50, .1), limit = c(0,.50))
```

***
> *Observation*: Looks like a low debt to income ration and a high score
will earn you a low borrower APR.

# 22B.Linear model for Credit score and borrower apr
```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x=CreditScoreRangeLower, y= borrower_apr),
       data=loan) +
  geom_point( position=position_jitter(h=0)) +
   scale_x_continuous(name= 'Credit Score', limits = c(400, 880)) +
   scale_y_continuous(name="Borrower APR",
                     breaks = seq(0, .50, .1), limit = c(0,.50)) + 
  geom_smooth(method='lm', color='yellow')

with(loan, cor.test(CreditScoreRangeLower, loan$borrower_apr, method='pearson'))
```

***
> *Observation*: The linear model shows that there is a negative relationship
between the variables of borrower APR and Credit Score.  The pearson
correlation coefficient is -0.4297, which indicates that there is a
weak negative correlation.

***

#22C. Credit score and apr with percentile and average.
```{r echo=FALSE,  message=FALSE, warning=FALSE}
ggplot(aes(x=CreditScoreRangeLower, y= borrower_apr),
       data=loan) +
  geom_point(alpha= 1/20, position=position_jitter(h=0), color = 'orange') +
   scale_x_continuous(name= 'Credit Score', limits = c(400, 880)) +
   scale_y_continuous(name="Borrower APR",
                     breaks = seq(0.00, .50, .1), limit = c(0.00,.50)) +
  
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat='summary', fun.y=quantile,
            fun.args = list(probs=.90),
            linetype = 2, color ='red') +
  geom_line(stat='summary', fun.y=quantile,
            fun.args = list(probs=.10),
            linetype = 2, color ='blue') 
 
```

***
> *Observation*: This plot shows that there is a a lot of variability with
Borrower APR.  There are more factors than just Credit Score that have
an effect on Borrower APR such as Debt-To-Income ratio. However,
having a high Credit Score does have an advantage as the lines show
that on average the lowest Borrower APR is given to the people with the
Highest Credit Scores.

***


# 23. Multivariate Analysis

> **Summary**: The features of interest in this data set are Borrower APR
and the variables/features that affect the APR positively or negatively.
For, example I want to know how the Credit Score, Prosper Score, and 
Prosper Rating affect the APR.  I noticed a lot of variability such as 
borrowers with a high Debt-to-income that received a low borrower APR.
However, on average if a borrower has a good credit score, the borrower
should expect to get a low Borrower APR.  
One interaction that I found interesting is in section #21.  I looked at 
this plot to consider if monthly payment has an effect on loan status. I 
thought that perhaps when a person has a high monthly payment they can 
over extend themselves causing them to default.  However, people with 
a low credit rating such as 'E' are trusted with low monthly payments 
but that is not enough to prevent them from defaulting on the loan.  
I understand that there are other factors such as Debt-to-Income Ratio
and employment status that come into play as far as a person's ability 
to repay a loan.

> *Models*:
The data set has a lot of variablity and thus there is a low negative\
correlation between Borrower APR and Credit Score.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

# 24. Final Plot #1: Original Loan Amount by Prosper Rating
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(aes(x=prosper_rating_a, y= LoanOriginalAmount), data = loan) +
  geom_boxplot(color = 'black', fill= '#099DD9') + 
  xlab('Prosper Rating (Descending Order)') +
  ylab('Original Loan Amount (US Dollars)') +
  
  ggtitle("Original Loan Amount by Prosper Rating")
```

> **Description for Final Plot #1**:This plot provides an answer to the question 
as to the benefits of having a good credit score 
(in this case a good Prosper Rating).  Clearly, having a Prosper Rating of 
'AA', 'A' or 'B' have a higher median Loan Amount and are even 
trusted with loan amounts above $ 30,000.  A Prosper Rating of 'C' 
and below all have lower median Loan amounts.

# 25. Final Plot #2: Borrower Apr vs. Credit Score
```{r echo=FALSE,  message=FALSE, warning=FALSE, Plot_Two}
ggplot(aes(x=CreditScoreRangeLower, y= borrower_apr),
       data=loan) +
  geom_point(alpha= 1/20, 
             position=position_jitter(h=0), 
             color = 'orange') +
   scale_x_continuous(name='Credit Score',
                      breaks = seq(400, 900, 50),
                      limits = c(400, 880)) +
   scale_y_continuous(name="Borrower APR",
                     breaks = seq(0, .50, .1), 
                     limit = c(0,.50)) +
  
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat='summary', fun.y=quantile,
            fun.args = list(probs=.90),
            linetype = 2, color ='red') +
  geom_line(stat='summary', fun.y=quantile,
            fun.args = list(probs=.10),
            linetype = 2, color ='blue') +
 
  ggtitle('Borrower APR vs. Credit Score\n 90th
          Percentile, Mean, and 10th percentile')
```



> **Description for Final Plot #2**: This plot shows that there is a 
lot of variability with Borrower APR.  There are more factors than 
just Credit Score that have an effect on Borrower APR such as 
Debt-To-Income ratio. However,having a high Credit Score does 
have an advantage as the lines show that on average the lowest 
Borrower APR is given to the people with the Highest Credit Scores.
Furthermore, the blue line(10th perentile) shows that about a 750 credit
score those borrowers received an APR below 10%.  

# 26. Final Plot #3: Monthly Loan Payment vs Loan Status,\
colored by prosper rating
```{r echo=FALSE, fig.width=11, message=FALSE, warning=FALSE, Plot_Three}
p1 <- ggplot(aes(x= loan_status, y = MonthlyLoanPayment,
                 color=prosper_rating_a ), data=loan) + 
      geom_boxplot() + 
      theme_dark() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_color_brewer(type='qual') +
      xlab('Loan Status') +
      ylab('Monthly Loan Payment(Dollars)') +
      ggtitle("Monthly Payment vs. Loan Status: Colored by Prosper Rating")

p1

by(loan$prosper_rating_a, loan$loan_status, summary)
```

> **Description for Final Plot #3**: 83% of the loans are either
current or have been completed, 16.9% defaulted on their loan
(This includes Charge-offs which is equal to 10.5%). This plot 
allows us to use a granulirized approach to visualize loan status. 
For example, only 3.3% of borrowers who were Chargedoff had a "AA"
Prosper Rating.  Whereas, 20.48% of borrowers who were Chargedoff had 
an 'HR' Prosper Rating.  Lastly, this plot helps us visualize how 
montly payment is part of the loan status equation.  As expected the median
monthly payment for an 'AA' Prosper Score is higher the median monthly payment
for an 'E' score.  This helps to tell the story of how much risk Prosper Loans 
is willing to take on borrowers.  Borrowers with a high Prosper Rating 
are trusted with higher montly payments and are more likely to complete
their loans.


------

# 27. Reflection

> **Reflection**: In the introduction over-arching questions were posed:
Questions: What kinds of loan terms do people with excellent, good,
fair, poor, and bad credit receive? What are the advantages of having good 
credit?
To answer this question I created multiple plots to display and understand the 
data.  The most important factors are: What is my credit score/rating?
How much can I borrow(Loan Amount)?, How much will it cost to borrow the 
money(Borrower APR)? In the data I found that there is not one factor that can 
predict a Borrower APR. For example,  there is substantial variability when 
comparing Borrower APR vs. Credit Score, however, on average as the Credit 
Score increases the Borrower APR decreases and the amount of the loan also 
increases.  Meaning that the advantages of having a good credit are that it 
costs less to borrow money and there is trust with higher loan amounts. I found
surprising that borrowers that receive a prosper rating of 'AA' and 'A' are 
considered a risk to the lender since there are some who default on their loan
and even get Charged off. 
One of the main struggles was the complexity of the dataset since it contains 82 
variables.  A lot of exploration was necessary to understand which are the most
relevant variables.  In fact, I found that a deeper analysis needs to be done to 
find which variables have the most impact on Borrower APR.  A multiple regression 
model would be the next steps in this exploration with variables such as 
Debt-To-Income-Ratio, Prosper Score, Prosper Rating, Credit Score, 
Employment Duration,Homeownership could be used to predict Borrower APR 
and Loan Amount. Lastly, there was no time-series plots included in the analysis.
For further study I would like to look at how loan amounts varied over time.



